---
title: "A few swiss indices by cantons"
author: "Lynda Metref"
date: "16 January 2016"
output: html_document
---

Downloading data from bfs website at `http://www.bfs.admin.ch/xmlns/opendata/je-f-21.03.02.ods`. Remove comments at end of sheet.

```{r}
library(readODS)
library(dplyr)
library(tidyr)
library(XML)


#download.file("http://www.bfs.admin.ch/xmlns/opendata/je-e-21.03.02.ods","canton-ind.ods")
dataStr <- read.ods(file = "canton-ind.ods", sheet = 1)
comments <- dataStr[89:nrow(dataStr),1]
dataStr <- dataStr[1:86,]
```


- transpose: each canton is a row because each canton is an observation and the indcies are the values.
```{r}
dataStr=as.data.frame(t(dataStr),stringsAsFactors = FALSE)
```


Canton names have not been imported by read.ods for some reason... Need to be automatised somehow...
```{r}
dataStr <- cbind(c("","Années","Suisse",'CH-ZH','CH-BE','CH-LU','CH-UR','CH-SZ','CH-OW','CH-NW','CH-GL','CH-ZG','CH-FR','CH-SO','CH-BS','CH-BL','CH-SH','CH-AR','CH-AI','CH-SG','CH-GR','CH-AG','CH-TG','CH-TI','CH-VD','CH-VS','CH-NE','CH-GE','CH-JU'),dataStr)
```

Lots of wrangling to do: 

- remove unused year
```{r}
dataStr <- dataStr[-2,]
```


- remove first row and rename colums with it
```{r}
colnames(dataStr) <- dataStr[1,]
dataStr <- dataStr[-1,]
```

- remove first column and rename rows with it
```{r}
rownames(dataStr) <- dataStr[,1]
dataStr <- dataStr[,-1]
```

- removing all blank lines (when first cell is blank the whole line is)
```{r}
dataStr <- select(dataStr,-matches("^[[:punct:][[:digit:]]+"))
dataStr <- dataStr[,-12]
```

- remove parenthesis and convert string to numeric
```{r}
dataNum <- mapply(gsub,"[\\'|\\(|\\)|X|\\*]","",dataStr)
dataNum<-as.data.frame(apply(dataNum,2,as.numeric))
colnames(dataNum) <- colnames(dataStr)
rownames(dataNum) <- rownames(dataStr)
```

- Extract population data first column must be the location for the gvis map
- multiply resitdent to have the right number
- beautifuling a bit the titles
```{r}
population <- dataNum[c(2,3,4,7)]
populationCH <- population[1,]
population <- population[-1,]
population <- cbind(canton=rownames(population),population)
population[,2] <- population[,"Residents in 1000"]*1000
colnames(population) <- c("canton","Residents in the canton","Residents per km^2","Change in % between 2010 and 2012","Foreign nationals in %")

save(population,populationCH,file="population.rda")
```



