---
title: "A few swiss indices by cantons"
author: "Lynda Metref"
date: "16 January 2016"
output: html_document
---

Downloading data from bfs website at `http://www.bfs.admin.ch/xmlns/opendata/je-f-21.03.02.ods`. Data are in french. Remove comment at end of sheet. Convert to tdl for easier wrangling with dplyr.

```{r}
library(readODS)
library(dplyr)
library(tidyr)
library(XML)


#download.file("http://www.bfs.admin.ch/xmlns/opendata/je-f-21.03.02.ods","canton-ind.ods")
dataStr <- read.ods(file = "canton-ind.ods", sheet = 1)
comments <- dataStr[89:nrow(dataStr),1]
dataStr <- dataStr[1:86,]
```


- transpose: each canton is a row because each canton is an observation and the indcies are the values.
```{r}
dataStr=as.data.frame(t(dataStr),stringsAsFactors = FALSE)
```


Canton names have not been imported by read.ods for some reason... Need to be automatised somehow...
```{r}
dataStr <- cbind(c("","Années","Suisse",'ZH','BE','LU','UR','SZ','OW','NW','GL','ZG','FR','SO','BS','BL','SH','AR','AI','SG','GR','AG','TG','TI','VD','VS','NE','GE','JU'),dataStr)
```

# Trying to extract canton names since they have not been imported by readODS
# unzip("canton-ind.ods", files= "content.xml")
# Extract XML element table:table-row table:style-name="ro1" which is the row containing the information

Lots of wrangling to do: 

-remove unused year
```{r}
dataStr <- dataStr[-2,]
```


- remove first row and rename colums with it
```{r}
colnames(dataStr) <- dataStr[1,]
dataStr <- dataStr[-1,]
```

- remove first column and rename rows with it
```{r}
rownames(dataStr) <- dataStr[,1]
dataStr <- dataStr[,-1]
```

- removing all blank lines (when first cell is blank the whole line is)
```{r}
dataStr <- select(dataStr,-matches("^[[:punct:][[:digit:]]+"))
dataStr <- dataStr[,-12]
```

- remove parenthesis and convert string to numeric
```{r}
dataNum <- mapply(gsub,"[\\'|\\(|\\)|X|\\*]","",dataStr)
dataNum<-as.data.frame(apply(dataNum,2,as.numeric))
colnames(dataNum) <- colnames(dataStr)
rownames(dataNum) <- rownames(dataStr)


```

- Extract population data
```{r}
population <- dataNum[2:7]
populationCH <- population[1,]
population <- population[-1,]
save(population,populationCH,file="population.rda")
```



